Follow Up Samples

### 2016-05-02 Baxter_followUps_2016.Rmd
added change in model by tumor stage

```{r otu_model, echo=FALSE, cache=TRUE, fig.height=5, fig.width=6}
# Make training set with samples withouth follow ups
train <- merge(metaI, shared, by.x='sample', by.y='Group')
train <- train[-match(metaF$initial, train$sample),]
train <- train[,c('lesion',colnames(train)[grep('Otu[0123456789]', colnames(train))])]

@@

```{r dif_otu, echo=FALSE, fig.height=6, fig.width=6}
plot(0, type='n', xlim=c(0.5,4.5), ylim=c(0,1.1), ylab='Probability of lesion (OTU model)', xaxt='n', xlab='')
mtext(text=c('Adenoma','Cancer'), at=c(1.5,3.5), side=1, line=2, font=2)
axis(1, at=c(1,2,3,4), labels=c('Baseline','Follow Up','Baseline','Follow Up'))

@@

@@

>>>>

```{r otu_model, echo=FALSE, cache=TRUE, fig.height=5, fig.width=6}
# Make training set with samples withouth follow ups
train <- merge(metaI, shared, by.x='sample', by.y='Group')
train <- train[-match(metaF$initial, train$sample),]
train <- train[metaI$dx!='adenoma',c('lesion',colnames(train)[grep('Otu[0123456789]', colnames(train))])]

@@

```{r dif_otu, echo=FALSE, fig.height=6, fig.width=6}

plot(0, type='n', xlim=c(0.5,4.5), ylim=c(0,1.1), ylab='Probability of lesion (OTU model)', xaxt='n', xlab='')
mtext(text=c('Adenoma','Cancer'), at=c(1.5,3.5), side=1, line=2, font=2)
axis(1, at=c(1,2,3,4), labels=c('Baseline','Follow Up','Baseline','Follow Up'))

@@

points(rep(3,26), initial_probs[metaF$dx=='cancer'], pch=c(21,22,23,24,25)[metaF$stage], bg='red', cex=0.8, lwd=0.5)
points(rep(4,26), follow_probs[metaF$dx=='cancer'], pch=c(21,22,23,24,25)[metaF$stage], bg='red', cex=0.8, lwd=0.5)
segments(3, initial_probs[metaF$dx=='cancer'], 4, follow_probs[metaF$dx=='cancer'], col='red')
canc_p <- t.test(initial_probs[metaF$dx=='cancer'], follow_probs[metaF$dx=='cancer'], paired=T, alternative='greater')$p.value
text(3.5, 1, sprintf('p=%.1g',canc_p))

@@

```{r diff_by_stage, echo=FALSE, cache=TRUE, fig.height=5, fig.width=6}
# Make training set with samples withouth follow ups
train <- merge(metaI, shared, by.x='sample', by.y='Group')
train <- train[-match(metaF$initial, train$sample),]
train <- train[, c('lesion',colnames(train)[grep('Otu[0123456789]', colnames(train))])]

set.seed(050216)
rf_otu <- AUCRF(lesion~., data=train, pdel=0.1, ntree=500, ranking='MDA')
rf_opt <- rf_otu$RFopt
#plot(rf_otu)

layout(matrix(c(1,2,3,4), nrow=1))
par(mar=c(5,3,1,0.5))
#non advanced
plot(0, type='n', xlim=c(0.75,2.25), ylim=c(0,1.1), ylab='Probability of lesion (OTU model)', xaxt='n', xlab='')
mtext(text='Nonadvanced\nAdenoma', at=c(1.5), side=1, line=4, font=2)
axis(1, at=c(1,2,3,4), labels=c('Baseline','Follow Up','Baseline','Follow Up'))
stripchart(at=1, initial_probs[metaF$Dx=='Adenoma'], pch=21, bg='orange', cex=1.2, lwd=0.5, vertical=T, add=T)
stripchart(at=2, follow_probs[metaF$Dx=='Adenoma'], pch=21, bg='orange', cex=1.2, lwd=0.5, vertical=T, add=T)
segments(1, initial_probs[metaF$Dx=='Adenoma'], 2, follow_probs[metaF$Dx=='Adenoma'], col='orange')
ade_p<-t.test(initial_probs[metaF$Dx=='Adenoma'], follow_probs[metaF$Dx=='Adenoma'], paired=T, alternative='greater')$p.value
text(1.5, 1, sprintf('p=%.1g',ade_p))

#advanced
plot(0, type='n', xlim=c(0.75,2.25), ylim=c(0,1.1), ylab='Probability of lesion (OTU model)', xaxt='n', xlab='')
mtext(text='Advanced\nAdenoma', at=c(1.5), side=1, line=4, font=2)
axis(1, at=c(1,2,3,4), labels=c('Baseline','Follow Up','Baseline','Follow Up'))
stripchart(at=1, initial_probs[metaF$Dx=='adv Adenoma'], pch=21, bg='orange', cex=1.2, lwd=0.5, vertical=T, add=T)
stripchart(at=2, follow_probs[metaF$Dx=='adv Adenoma'], pch=21, bg='orange', cex=1.2, lwd=0.5, vertical=T, add=T)
segments(1, initial_probs[metaF$Dx=='adv Adenoma'], 2, follow_probs[metaF$Dx=='adv Adenoma'], col='orange')
ade_p<-t.test(initial_probs[metaF$Dx=='adv Adenoma'], follow_probs[metaF$Dx=='adv Adenoma'], paired=T, alternative='greater')$p.value
text(1.5, 1, sprintf('p=%.1g',ade_p))

#Early (stage 1 or 2)
plot(0, type='n', xlim=c(0.75,2.25), ylim=c(0,1.1), ylab='Probability of lesion (OTU model)', xaxt='n', xlab='')
mtext(text='Cancer\nStage I or II', at=c(1.5), side=1, line=4, font=2)
axis(1, at=c(1,2,3,4), labels=c('Baseline','Follow Up','Baseline','Follow Up'))
stripchart(at=1, initial_probs[metaF$stage==1 | metaF$stage==2], pch=21, bg='red', cex=1.2, lwd=0.5, vertical=T, add=T)
stripchart(at=2, follow_probs[metaF$stage==1 | metaF$stage==2], pch=21, bg='red', cex=1.2, lwd=0.5, vertical=T, add=T)
segments(1, initial_probs[metaF$stage==1 | metaF$stage==2], 2, follow_probs[metaF$stage==1 | metaF$stage==2], col='red')
ade_p<-t.test(initial_probs[metaF$stage==1 | metaF$stage==2], follow_probs[metaF$stage==1 | metaF$stage==2], paired=T, alternative='greater')$p.value
text(1.5, 1, sprintf('p=%.1g',ade_p))

#Early (stage 3 or 4)
plot(0, type='n', xlim=c(0.75,2.25), ylim=c(0,1.1), ylab='Probability of lesion (OTU model)', xaxt='n', xlab='')
mtext(text='Cancer\nStage III or IV', at=c(1.5), side=1, line=4, font=2)
axis(1, at=c(1,2,3,4), labels=c('Baseline','Follow Up','Baseline','Follow Up'))
stripchart(at=1, initial_probs[metaF$stage==3 | metaF$stage==4], pch=21, bg='red', cex=1.2, lwd=0.5, vertical=T, add=T)
stripchart(at=2, follow_probs[metaF$stage==3 | metaF$stage==4], pch=21, bg='red', cex=1.2, lwd=0.5, vertical=T, add=T)
segments(1, initial_probs[metaF$stage==3 | metaF$stage==4], 2, follow_probs[metaF$stage==3 | metaF$stage==4], col='red')
ade_p<-t.test(initial_probs[metaF$stage==3 | metaF$stage==4], follow_probs[metaF$stage==3 | metaF$stage==4], paired=T, alternative='greater')$p.value
text(1.5, 1, sprintf('p=%.1g',ade_p))
